---
title: "EnsembleModels"
author: "Lisa"
date: "March 12, 2015"
output: pdf_document
---

# Introduction

I applied a package **caretEnsemble** for making ensembles of caret models. It written by *Zach Mayer* and his gitbut website is <https://github.com/zachmayer/caretEnsemble>. The **caretEnsemble** includes 2 different algorithms for combining models:

1. Greedy stepwise ensembles (returns a weight for each model), using **caretEnsemble**.
2. Stacks of caret models, using **caretStack**. The stacking algorithm simply builds a second caret model on top of the existing models (using their predictions as input).



###Resourse:

* Binary classification model with caretEnsemble on gist of *Zach Mayer* <https://gist.github.com/zachmayer/5179418/>
* A brief introduction to caretEnsemble <http://cran.r-project.org/web/packages/caretEnsemble/vignettes/caretEnsemble-intro.html>


```{r setup,include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```


##Load libraries

```{r,echo=FALSE,message=FALSE}

#Setup
rm(list = ls(all = TRUE)) 
gc(reset=TRUE) #maximum space for using
 
#Libraries
library(devtools)
#install_github('caretEnsemble', 'zachmayer')  #Install zach's caretEnsemble package
library(caretEnsemble)
#devtools::install_github('topepo/caret/pkg/caret') #Install the development version from GitHub
library(caret)
library ( kernlab ) 
library(gbm)
library(ggplot2)
library(doParallel)
library(plyr)
library(dplyr)

```

I deleted missing values (*NA*) in the dataset and added levels for response as "good" and "bad".

```{r}
load("data_17_Feb.RData")
data=data[,2:56]

# delete 2557 missing values in "compt_size"
sum(is.na(data$compt_size))
data=na.omit(data)

dim(data)
sum(is.na(data)) #check missing value

#make levels for response
levels(data$class)=c("good","bad")
str(data)
prop.table(table(data$class)) #proportions of our outcome variable
```

Our data isn't perfectly balanced but it certainly isn't skewed or considered a rare event.

I added levels for response, because when you ask for class probabilities, model predictions are a data frame with separate columns for each class/level. If *class* doesn't have levels, `data.frame` converts them to valid names, which creates a problem because a different (but valid) name of levels.


## Data splitting for training and testing

```{r,echo=FALSE}
set.seed(100)
#Spliting data as training and test set. 75% percentage of data in the training set.
Train <- createDataPartition(y =data$class,p = 0.75,list = FALSE)  
train <- data[Train,]
test <- data[-Train,]
dim(train)
dim(test)

```



```{r,echo=FALSE,eval=FALSE}
#Ignore this subset
set.seed(10)
inTrain<-createDataPartition(y =data$class,p = 0.1,list = FALSE)  
Data<-data[inTrain,]
dim(Data)

set.seed(100)
#Spliting data as training and test set. 90% percentage of data in the training set.
Train <- createDataPartition(y =Data$class,p = 0.75,list = FALSE)  
train <- Data[Train,]
test <- Data[-Train,]
dim(train)
dim(test)
```

# Model tuning

```{r,eval=FALSE}
#Setup CV Folds
#returnData=FALSE saves some space
folds=10
repeats=3
# Try using ROC (AUC) as metric since that is what caretEnsemble uses
#trainMetric = "ROC" # Classification, required for some two class analyses

myControl <- trainControl(method='cv', number=folds, repeats=repeats, 
                          returnResamp='none', classProbs=TRUE,
                          returnData=FALSE, savePredictions=TRUE, 
                          verboseIter=TRUE, allowParallel=TRUE,
                          summaryFunction=twoClassSummary,
                          # For caretEnsemble, make the resampling indexes all the same.
                          index=createMultiFolds(train$LOAN_CLASS, k=folds, times=repeats)
                          )
PP <- c('center', 'scale')


```




```{r,eval=FALSE}
set.seed(800)
detectCores()
registerDoParallel(48,cores=48)
getDoParWorkers()

model_list<- caretList(
  LOAN_CLASS~., data=train,
  trControl=myControl,
  methodList=c('blackboost', 'parRF'),
  tuneList=list(
    #gbm=caretModelSpec(method='gbm',
                       #tuneGrid=expand.grid(.n.trees=c(1:10)*10, .interaction.depth=1, .shrinkage = 0.1)
                       #),
    mlpweightdecay=caretModelSpec(method='mlpWeightDecay', trace=FALSE, preProcess=PP),
    earth=caretModelSpec(method='earth', preProcess=PP),
    glm=caretModelSpec(method='glm', preProcess=PP),
    svmRadial=caretModelSpec(method='svmRadial', preProcess=PP),
    knn=caretModelSpec(method='knn',preProcess=PP),
    #gam=caretModelSpec(method='gam', preProcess=PP),
    glmnet=caretModelSpec(method='glmnet', preProcess=PP)
    )
  )

```

```{r,eval=FALSE,echo=FALSE}
#Ignore these individual models
save(file="modellist",list="model_list")
model1 <- train(class~., data=train, method='gbm', trControl=myControl,
                tuneGrid=expand.grid(.n.trees=c(1:20)*10, .interaction.depth=c(1:5), .shrinkage = 0.1))
model2 <- train(class~., data=train, method='blackboost', trControl=myControl)
model3 <- train(class~., data=train, method='parRF', trControl=myControl)
model4 <- train(class~., data=train, method='mlpWeightDecay', trControl=myControl, trace=FALSE, preProcess=PP)
model6 <- train(class~., data=train, method='earth', trControl=myControl, preProcess=PP)
model7 <- train(class~., data=train, method='glm', trControl=myControl, preProcess=PP)
model8 <- train(class~., data=train, method='svmRadial', trControl=myControl, preProcess=PP)
model9 <- train(class~., data=train, method='gam', trControl=myControl, preProcess=PP)
model10 <- train(class~., data=train, method='glmnet', trControl=myControl, preProcess=PP)
```

#Benchmark models (candidate)


We considered 9 candidate models. Some are the least interpretable and most flexible, such as *boosted trees*, *Stochastic Gradient Boosting* or *support vector machines* which have a high likelihood of producing the most accurate results. Others are simpler models that are less opaque (e.g., not complete black boxes), such as *multivariate adaptive regression splines (MARS)* or *generalized additive models*. The model and argument value are in `names(model_list)` and `names(infor)`, respectively. The more detail method information such as tuning parameters can be explored on [caret model list](http://topepo.github.io/caret/modelList.html).



```{r,message=FALSE,warning=FALSE}
load("modellist3") #"model_list" is saved in "modellist3"

#Make a list of all the models
names(model_list) = sapply(model_list, function(x) x$method) #method names
infor = sapply(model_list, function(x) x$modelInfo$label) #label of methods
names(infor) = infor 
names(model_list)
names(infor)
sort(sapply(model_list, function(x) max(x$results$ROC))) #maximum accuracy of each method

```

#Make a greedy ensemble


```{r,message=FALSE,warning=FALSE}
#Make a greedy ensemble - currently can only use RMSE
greedy <- caretEnsemble(model_list, iter=1000L)
sort(greedy$weights, decreasing=TRUE)
#summary(greedy)
greedy$error
```

The greedy model relies 87\% on the **blackboost**, **glm** and **earth**, which makes sense as these are top three highest accuracte models on the training set. The ensemble’s AUC (area under the curve) on the training set is 0.87, which is about 1\% better than the best individual model (**blackboost**).

#Make a linear regression ensemble

```{r,warning=FALSE,message=FALSE}
#Make a linear regression ensemble
linear <- caretStack(model_list, method='glm', trControl=trainControl(method='cv'))
summary(linear$ens_model$finalModel)
linear$error
CF <- coef(linear$ens_model$finalModel)[-1] #coefficient of glm model
CF/sum(CF) #glm model weights
```

The linear model uses all of the models, and achieves an AUC of 0.81, which lower than the average individual model accuracy. I'm not sure if this is a failure of the stacking model, because accuracy becomes much higher later on in the test set prediction. 

Different from greedy model, The glm-weighted model weights relies most on only **blackboost**, and the weight is 0.87.


#Result (predict for test set)


```{r,eval=FALSE}
library(caTools)
model_preds <- lapply(model_list, predict, newdata=test, type='prob')
model_pred <- lapply(model_preds, function(x) x[,'good'])
model_pred <- data.frame(model_pred)
ENS_greedy <- predict(greedy, newdata=test)
model_pred$ENS_greedy=1-ENS_greedy
ENS_linear= predict(linear, newdata=test,type='prob')
model_pred$ENS_linear <- ENS_linear[,"good"]
```

model_pred$truth=test$class
allgood=model_pred[apply(model_pred[,c(1:9)]>=0.7,1,all) , ]  # patients only have triage variables
allbad=model_pred[apply(model_pred[,c(1:9)]<0.5,1,all) , ]  # patients only have triage variables
dim(allgood)
dim(allbad)
dim(allgood)[1]+dim(allbad)[1]

table(allgood$truth)
table(allbad$truth)

allgood=model_preds[apply(model_preds[,c(1:)]=="good",1,all) , ]  # patients only have triage variables
allbad=model_preds[apply(model_preds[,c(1:3)]=="bad",1,all) , ]  # patients only have triage variables
dim(allgood)
dim(allbad)

allgood

model_pred$gam=1-model_pred$gam

pred=rep("bad",3429)
pred[model_pred$gbm>.6]="good"
pred=factor(pred)
levels(pred)=c("bad","good")
table(pred,test$class)
con=confusionMatrix(pred, test$class)
confusionMatrix(predict(svm.tune1, test[,-1]), test$Diagnosis)

sum(model_pred$gbm>=.5)

```{r}
library(caTools)
load("pred")  #"model_pred" is saved in "pred"
sort(data.frame(colAUC(model_pred, test$class)))
```

The greedy model testing prediction still better than any individual model prediction, and has the highest accuracy 0.87.

Most predictions in test set are reasonable, except "linear regression ensemble", I mentioned before, that it has much higher testing accuracy 0.867 than training accuracy 0.81. *I've no idea why it happened*.